Arrows for Parallel Computations

### 2017-10-17 paper/content/benchmarks.tex
wip

The \ensuremath{\Conid{Par}} monad and GpH Haskell - in its multicore version (\cite{Marlow2009}) -  can be executed on a shared
memory machines only. Although GpH is available on distributed memory
clusters, and newer distributed memory Haskells such as HdpH exist,
current support of distributed memory in PArrows is limited to
Eden. We used the MPI backend of Eden in a distributed memory
setting. However, for shared memory Eden features a ``CP'' backend
that merely copies the memory blocks between distributed heaps. In
this mode, Eden still operates in the ``nothing shared'' setting, but
is adapted better to multicore machines. We label this version of Eden
in the plots as ``Eden~CP''.

>>>>

The \ensuremath{\Conid{Par}} monad and GpH Haskell -- in its multicore version (\cite{Marlow2009}) --  can be executed on a shared
memory machines only. Although GpH is available on distributed memory
clusters, and newer distributed memory Haskells such as HdpH exist,
current support of distributed memory in PArrows is limited to
Eden. We used the MPI backend of Eden in a distributed memory
setting. However, for shared memory Eden features a ``CP'' backend
that merely copies the memory blocks between distributed heaps. In
this mode, Eden still operates in the ``nothing shared'' setting, but
is adapted better to multicore machines. We label this version of Eden
in the plots as ``Eden~CP''.

### 2017-10-17 paper/content/parallelHaskells.tex
wip

\subsubsection{Glasgow parallel Haskell - GpH}
GpH \cite{Marlow2009,Trinder1998a} is one of the simplest ways to do parallel processing found in standard GHC.\footnote{The Multicore implementation of GpH is available on Hackage under \url{https://hackage.haskell.org/package/parallel-3.2.1.0}, compiler support is integrated in the stock GHC.} Besides some basic primitives (|par| \& |pseq| hints), it ships with parallel evaluation strategies for several types which can be applied with |using :: a -> Strategy a -> a|, which is exactly what is required for an implementation of |parEvalN|.

>>>>

\subsubsection{Glasgow parallel Haskell -- GpH}
GpH \cite{Marlow2009,Trinder1998a} is one of the simplest ways to do parallel processing found in standard GHC.\footnote{The Multicore implementation of GpH is available on Hackage under \url{https://hackage.haskell.org/package/parallel-3.2.1.0}, compiler support is integrated in the stock GHC.} Besides some basic primitives (|par| \& |pseq| hints), it ships with parallel evaluation strategies for several types which can be applied with |using :: a -> Strategy a -> a|, which is exactly what is required for an implementation of |parEvalN|.

### 2017-10-17 paper/content/utilityFunctions.tex
replaced "Multicore Haskell" with "GpH"
wip

\begin{figure}
\begin{code}
instance (NFData b, ArrowApply arr, ArrowChoice arr) => ArrowParallel arr a b () where
    parEvalN _ fs = parEvalN (hack fs) fs
                    where
                        hack :: (NFData b) => [arr a b] -> Conf b
                        hack _ = Conf rdeepseq
\end{code}
\caption{The actual default implementation of Multicore's |ArrowParallel|.}
\label{fig:actualNonConfigMulticore}
\end{figure}

@@

The |parEval2| skeleton is defined in Figure~\ref{fig:parEval2}. 
We start by transforming the |(a, c)| input into a two-element list |[Either a c]| by first tagging the two inputs with |Left| and |Right| and wrapping the right element in a singleton list with |return| so that we can combine them with |arr (uncurry (:))|. Next, we feed this list into a parallel arrow running on two instances of |f +++ g| as described above. After the calculation is finished, we convert the resulting |[Either b d]| into |([b], [d])| with |arr partitionEithers|. The two lists in this tuple contain only one element each by construction, so we can finally just convert the tuple to |(b, d)| in the last step.
\begin{figure}[h]
\begin{code}
parEval2 :: (ArrowChoice arr,
	ArrowParallel arr (Either a c) (Either b d) conf) =>
	conf -> arr a b -> arr c d -> arr (a, c) (b, d)
parEval2 conf f g =
	arr Left *** (arr Right >>> arr return) >>>
	arr (uncurry (:)) >>>
	parEvalN conf (replicate 2 (f +++ g)) >>>
	arr partitionEithers >>>
	arr head *** arr head
\end{code}
	\caption{Definition of |parEval2|.}
	\label{fig:parEval2}
\end{figure}
Furthermore, Fig.~\ref{fig:torus_example_rest} contains the ommited definitions of |prMMTr|, - which calculates $AB^T$ for two matrices $A$ and $B$, |splitMatrix| - which splits the a matrix into chunks, and lastly |matAdd|, that calculates $A + B$ for two matrices $A$ and $B$.
\begin{figure}[h]
\begin{code}
prMMTr m1 m2 = [[sum (zipWith (*) row col) | col <- m2 ] | row <- m1]

>>>>

\begin{figure}
\begin{code}
instance (NFData b, ArrowApply arr, ArrowChoice arr) => ArrowParallel arr a b () where
    parEvalN _ fs = parEvalN (hack fs) fs
                    where
                        hack :: (NFData b) => [arr a b] -> Conf b
                        hack _ = Conf rdeepseq
\end{code}
\caption{The actual default implementation of GpH's |ArrowParallel|.}
\label{fig:actualNonConfigMulticore}
\end{figure}

@@

The |parEval2| skeleton is defined in Figure~\ref{fig:parEval2}. 
We start by transforming the |(a, c)| input into a two-element list |[Either a c]| by first tagging the two inputs with |Left| and |Right| and wrapping the right element in a singleton list with |return| so that we can combine them with |arr (uncurry (:))|. Next, we feed this list into a parallel arrow running on two instances of |f +++ g| as described above. After the calculation is finished, we convert the resulting |[Either b d]| into |([b], [d])| with |arr partitionEithers|. The two lists in this tuple contain only one element each by construction, so we can finally just convert the tuple to |(b, d)| in the last step.
\begin{figure}[h]
\begin{code}
parEval2 :: (ArrowChoice arr,
	ArrowParallel arr (Either a c) (Either b d) conf) =>
	conf -> arr a b -> arr c d -> arr (a, c) (b, d)
parEval2 conf f g =
	arr Left *** (arr Right >>> arr return) >>>
	arr (uncurry (:)) >>>
	parEvalN conf (replicate 2 (f +++ g)) >>>
	arr partitionEithers >>>
	arr head *** arr head
\end{code}
	\caption{Definition of |parEval2|.}
	\label{fig:parEval2}
\end{figure}
Furthermore, Fig.~\ref{fig:torus_example_rest} contains the ommited definitions of |prMMTr| (which calculates $AB^T$ for two matrices $A$ and $B$), |splitMatrix| (which splits the a matrix into chunks), and lastly |matAdd|, that calculates $A + B$ for two matrices $A$ and $B$.
\begin{figure}[h]
\begin{code}
prMMTr m1 m2 = [[sum (zipWith (*) row col) | col <- m2 ] | row <- m1]
